\begin{frame}[allowframebreaks]
                \frametitle<presentation>{For Further Reading}

                \begin{thebibliography}{10}

                                \beamertemplatebookbibitems
                                % Start with overview books.
                                \bibitem[{Mitchell (2007)}]{mitchell07fusion}
                                {\sc Mitchell, H.B.}
                                \newblock {\em Multi-Sensor Data Fusion: An Introduction\/}
                                \newblock Springer, 1st edn,  2007.

                                \beamertemplatearticlebibitems
                                % Followed by interesting articles. Keep the list short. 

                                \bibitem[{Marchetti {\em et~al.\/} (2006)}]{MaGr06}
                                {\sc Marchetti, L., Grisetti, G. \& Iocchi, L.}
                                \newblock A comparative analysis of particle filter based localization methods.
                                \newblock {\em Proc.of RoboCup Symposium\/}, 2006.
                                
                                \bibitem[{Marchetti {\em et~al.\/} (2008)}]{marchetti08reliability}
                                {\sc Marchetti, L., Nobili, D. \& Iocchi, L.} (2008).
                                \newblock Improving tracking by integrating reliability of multiple sources.
                                \newblock {\em Conference on Information Fusion 11th International\/}, 2008.

                                \bibitem[{Makarenko {\em et~al.\/} (2004)}]{MakarenkoBrooks2004}
                                {\sc Makarenko, A., Brooks, A., Williams, S., Whyte, D.H. \& Grocholsky, B.}
                                \newblock A decentralized architecture for active sensor networks.
                                \newblock 2004

                                \bibitem[{Gutmann {\em et~al.\/} (2001)}]{Gutmann01}
                                {\sc Dietl~M., N.B., Gutmann J.-S.}
                                \newblock Cooperative sensing in dynamic environments.
                                \newblock {\em IEEE/RSJ International Conference on Intelligent Robots and Systems\/}, 2001.

                                \bibitem[{Rosencrantz {\em et~al.\/} (2003)}]{Rosencrantz2003}
                                {\sc Rosencrantz, M., Gordon, G. \& Thrun, S.}
                                \newblock Locating moving entities in dynamic indoor environments with teams of mobile robots.
                                \newblock {\em Second International Joint Conference on Autonomous Agents and Multiagent Systems\/}, 2001.

                                \bibitem[{Rogova (2004)}]{Rogova04}
                                {\sc Rogova, G.L.}
                                \newblock Reliability in information fusion: literature survey.
                                \newblock {\em 7th Intl. Conference on Information fusion\/}, 2004.

                                \bibitem[{Ceruti {\em et~al.\/} (2006)}]{ceruti06}
                                {\sc Ceruti, M., Das, S., Ashenfelter, A., Raven, G., Brooks, R., Sudit, M., Chen, G. \& Wright, E.}
                                \newblock Pedigree information for enhanced situation and threat assessment.
                                \newblock {\em 9th International Conference on\/}, 2006.

                                \bibitem[{Wright \& Laskey (2006)}]{Wright06}
                                {\sc Wright, E. \& Laskey, K.}
                                \newblock Credibility models for multi-source fusion.
                                \newblock {\em 9th International Conference on Information Fusion\/}, 2006.

                                \bibitem[{Yan (2007)}]{yan07}
                                {\sc Yan, W.}
                                \newblock Fusion in multi-criterion feature ranking.
                                \newblock {\em 10th International Conference on Information Fusion\/}, 2007.

                \end{thebibliography}
        \end{frame}
