%%
%% VERSION HISTORY
%%    22 May 2006 - John Papandriopoulos - Original version
%%    12 Jul 2007 - John Papandriopoulos - Converted into template
%%
\begin{acknowledgements}
Foremost, I would like to express my sincere gratitude to my two supervisors Prof. Philippe Bidaud and Prof. Vincent Padois for giving me the opportunity to be part of the \textit{Institut des syst\`emes intelligents et de robotique} (ISIR), for their continuous support during my Ph.D study and research, for their patience, motivation, enthusiasm, trust and immense knowledge. Their guidance made it possible to achieve the goals fixed for my thesis and also to write this report. \\
% it was impossible to reach the goal
%it has been a honour to be her their Ph.D. student.
%
%me during all the time of research and writing of this thesis and i could not have imagined having better advisors and mentors.  \\

Additionally, i would like to thank my committee members Dr. Olivier Stasse, Prof. Sami Haddadin, Prof. Guillaume Morel, Dr. Andrea Del Prete and Dr. Yvan Measson for their interest in my work. \\

I am also thankful to my colleagues Zhanat Kappassov, Ziad Zamzami, Mohammed Fnadi, Pauline Maurice, Ryan Lober, Mingxing Liu, Darwin Lau, Aurélien Ibanez, Philipp Schlehuner and Jorhabib Eljaik Gomez for their friendship, support, and for creating a cordial working environment. I would like also to thank Sovannara Hak, Antoine Hoarau and Jimmy Da Silva for their valuable help and technical assistance for the experimentations performed on the KUKA LWR4 robot. \\

Finally, in the pursuit of this Ph.D, the members of my family have been of great importance to me. More specifically, i would like to thank my mother Assia Meguenani and father Saadoun Meguenani for their unconditional love and moral and emotional support throughout my life and for making of me the person i am today. 

%Finally, nobody has been more important to me in the pursuit of this Ph.D than the members of my family. More specifically, i would like to thank my mother Assia Meguenani and father Saadoun Meguenani for their unconditional love and moral and emotional support throughout my life and for making of me the man i am today.
%
%Three year PhD experience is unforgettable and enjoyable for me. I am happy that I have learned many things concerning my research topics, the way to work, and the skills to live abroad. At the end of this journey, I would like to sincerely thank all the people who have been in my life and supporting me.
%
%First of all, I would like to thank my supervisor Prof. Philippe Bidaud for giving the opportunity to study at Pierre and Marie Currie University in France. I cherish this studying opportunity abroad. Not only because I can understand French customs and cultures, but also I can work in the institute for intelligent systems and robotics (ISIR), one of most famous robotic laboratory in the world. I would also like to thank Prof. Vincent Padois, who have provided me with his valuable time, patience and experience whenever it was needed. In addition to his guidance towards my PhD research, Vincent is like a friend. I also thank him for his nurturing to prepare me for the future with valuable life attitudes. 
%
%Secondly, appreciation goes to my colleagues in ISIR. I would like to thank Mingxing Liu and Darwin Lau for their inspiring discussions on robotics and their insightful comments on my research work. I would also like to thank Joseph Sainili, Sovannara Hak, Aurélien Ibanez, Pauline Maurice, Anis Meguenani, Ryan Lober, Tianming Lu, and Munan Yin for their ongoing support and advice both in and out of my research.
%
%
%Finally, I must acknowledge and thank my family. Thanks to my wife Dongya Liu for being with me and encouraging me during this challenging journey. I would also like to thank my parents, Zhizhong Tan and Yonghua Yang, for their everlasting love and support throughout my life. 
\end{acknowledgements}
